---
title: "R Bridge Course Final Project"
author: "MiaC"
date: "January 10, 2019"
output: html_document
---
## Question for analysis: 
### Which has greater influence on students' first year GPA in college? High school GPA or SAT score?
```{r}
# Read FirstYearGPA.csv from github
gpa <- read.csv("https://raw.githubusercontent.com/miachen410/FirstYearGPA/master/FirstYearGPA.csv")
```
## 1. Data Exploration
```{r}
# Explore the dataset by looking at the first few rows
head(gpa)

# Identify the types of variables
str(gpa)

# Generate a summary statistics including means, medians, quartiles, minimum and maximum
summary(gpa)
```
All variables are integer or numeric.
The average GPA is 3.096 and median GPA is 3.150.
The average High School GPA is 3.453, median GPA is 3.5.
Percentage of male is 46.58%.
Percentage of first generation college student is 11.42%.

```{r}
# Count numbers of male versus female students
table(gpa$Male)
```
There are 102 male students and 117 female students.

### 2. Data wrangling
```{r}
library(dplyr)
```
```{r}
# Add a new column called SAT, which combines SATV and SATM
gpa <- mutate(gpa, SAT = SATV + SATM)

# Filter male students
# Display only variables X, GPA, HSGPA, SAT, FirstGen, White
# Assign to new data frame males
males <- gpa %>% filter(Male == 1) %>% select(X, GPA, HSGPA, SAT, FirstGen, White)

# Show first 6 rows of males
head(males)
```

### 3. Graphics
```{r}
# Scatter Plot to show relationship between HSGPA and GPA in males
plot(males$HSGPA, males$GPA, main = "High School GPA v.s. First Year College GPA of Males", xlab = "High School GPA", ylab = "College GPA")
```
```{r}
# Histogram to show the distribution of GPA distribution among males
hist(males$GPA, main = "First Year College GPA Distribution among Males", xlab = "GPA")
```
```{r}
boxplot(GPA~Male, data = gpa, names = c("Female", "Male"), main = "First Year College GPA of Females and Males")
```
```{r}
library(ggplot2)
```
```{r}
# Scatter plot from ggplot2 package, showing relationship between SAT and college GPA for male representing by blue dots, and female representing by black dots
a <- ggplot(gpa, aes(x=SAT, y=GPA, color=Male)) + geom_point() + ggtitle("SAT v.s. College GPA for Male and Female")
a
```

```{r}
b <- ggplot(gpa, aes(x=HSGPA, y=GPA, color=Male)) + geom_point() + ggtitle("High School GPA v.s. College GPA for Male and Female")
b
```

## Observation:
### High school GPA seems to be more correlated with first year college GPA than SAT score.
